# Lasso with center

```{r, echo = FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE, echo = FALSE)
```


```{r}
library(patchwork)
library(ggpubr)
library(raster)
library(glmnet)
library(Hmisc)
source("../code/R/helper-functions.R")
```

```{r}
# inspect error plots
# inspect coef plots
# inspect prediction plots
# length(ids$train$Training060)
# length(ids$test$Testing060)
```
```{r}
path_to_model_folder <- "../results/CV-lasso/test-lasso-center"
```


## Model

We fit a LASSO model on the precipitation and SST data.
But in this case we standardize in the glmnet function.

## Error plots

<!-- ```{r} -->
<!-- err_mat <- readRDS("../results/CV-lasso/cv-lasso-og-data-16-06-22/err-mat.rds") -->
<!-- #err_mat <- sqrt(err_mat) -->
<!-- err_mat <- err_mat[1:60,] -->
<!-- ``` -->

<!-- ```{r} -->
<!-- lambda_id <- c(1:60) -->
<!-- mse <- apply(err_mat, 1, mean) -->
<!-- b <- mse -->
<!-- bplus <- apply(err_mat, 1, max) -->
<!-- bmin <- apply(err_mat, 1, min) -->
<!-- errbar(lambda_id, mse, b+bplus, b-bmin) -->
<!-- ``` -->
```{r}
err_mat <- readRDS(paste0(path_to_model_folder, "/err-mat.rds"))
lambdas <- readRDS(paste0(path_to_model_folder, "/lambda-vec.rds"))
```


```{r err-bar-plot-lasso-stand, fig.cap="Mean squared error of the 5-fold blocked cross validation for a range of lambda values on the log scale. The points in the middle represent the average MSE for the respective lambda, the errorbars give the MSE +/- one standard deviation. The dotted line shows the lambda for which minimum MSE was obtained."}
err_bars_plot <- readRDS(paste0(path_to_model_folder, "/err-mat-plots/err-bars-plot.rds"))
err_bars_plot
a <- ggplot_build(err_bars_plot)
```
\@ref(fig:err-bar-plot-lasso-stand) shows the results from the 5 fold-CV plotted for each lambda.
The lambdas are given on the log scale.
The upper and lower bars indicate mean MSE +/- one standard deviation from the
mean MSE. 
For values of log $\lambda$ smaller than 0 the differences in MSE become huge.

```{r}
#err_bars_plot + xlim(0,3) + ylim(-3000,3000)
```


<!-- ```{r} -->
<!-- # which.min(b) -->
<!-- # bs <- b[40:100] -->
<!-- # bbpluss <- (b+bplus)[40:100] -->
<!-- # bbmins <- (b-bmin)[40:100] -->
<!-- # (bs > bbpluss) | (bs < bbmins) -->
<!-- bmin <- which.min(b) -->
<!-- tv <- (b+bplus)[bmin] -->
<!-- tz <- b+bplus -->
<!-- tz <- tz[(bmin+1):100] -->
<!-- w <- tv < tz -->
<!-- #bmin is 40 -->
<!-- #w ist bei 45 true -->
<!-- l1se <- 45 -->
<!-- ``` -->



```{r err-fold-lasso-stand, fig.cap="MSE of the CV for the different lambda values on the a log scale. The red dotted line shows the lambda for which minimum MSE was obtained."}
p1 <- readRDS(paste0(path_to_model_folder, "/err-mat-plots/new-err-plot-fold-1.rds"))
p2 <- readRDS(paste0(path_to_model_folder, "/err-mat-plots/new-err-plot-fold-2.rds"))
p3 <- readRDS(paste0(path_to_model_folder, "/err-mat-plots/new-err-plot-fold-3.rds"))
p4 <- readRDS(paste0(path_to_model_folder, "/err-mat-plots/new-err-plot-fold-4.rds"))
p5 <- readRDS(paste0(path_to_model_folder, "/err-mat-plots/new-err-plot-fold-5.rds"))

p1 + p2 + p3 + p4 + p5

```
The errors in fold 5 are the ones that explode after lambda 0.


```{r}
err_mat_p_list <- list(p1,p2,p3,p4,p5)
err_mat_resc_list <- replot_err_mat(err_mat_p_list)
```

```{r err-fold-lasso-stand-replot, fig.cap="MSE of the CV for the different lambda values on the a log scale. The red dotted line shows the lambda for which minimum MSE was obtained. See @\ref(fig:err-folg-lasso-og), but this time the y-axis has the same range for all plots."}
err_mat_resc_list[[1]] + err_mat_resc_list[[2]] + err_mat_resc_list[[3]] + err_mat_resc_list[[4]] + err_mat_resc_list[[5]]
```
This plot emphasizes the findings from the plot above


Below we can see the minimum MSE for each fold.
```{r}
apply(err_mat, 2, min)
```
And which lambda in the lambda vector resulted in the lowest prediction error on the folds´ test set.

```{r}
apply(err_mat, 2, function(x) lambdas[which.min(x)])
```

## Coefficient plots

```{r coef-plot-lasso-stand, fig.cap="Coefficient map plot for the different folds. Longitude and Latitude on the x and y-axis respectively. Positive values are coloured in blue, negative values in red."}
pc1 <- readRDS(paste0(path_to_model_folder, "/coef-plots/coef-plot-fold-1.rds"))
pc2 <- readRDS(paste0(path_to_model_folder, "/coef-plots/coef-plot-fold-2.rds"))
pc3 <- readRDS(paste0(path_to_model_folder, "/coef-plots/coef-plot-fold-3.rds"))
pc4 <- readRDS(paste0(path_to_model_folder, "/coef-plots/coef-plot-fold-4.rds"))
pc5 <- readRDS(paste0(path_to_model_folder, "/coef-plots/coef-plot-fold-5.rds"))

l <- list(pc1,pc2,pc3,pc4,pc5)
ggarrange(plotlist = l, ncol = 3, nrow = 2,
         common.legend = TRUE)
```


## Inspect predictions from each fold

Following we inspect the folds precipitation time series and
the predictions made by the model.

```{r}
pred_plot_1 <- readRDS(paste0(path_to_model_folder, "/pred-plots/pred-plot-fold-1.rds"))
#pred_plot_1
mse_1 <- get_mse_from_pred_plot(pred_plot_1)
```
```{r}
pred_plot_2 <- readRDS(paste0(path_to_model_folder, "/pred-plots/pred-plot-fold-2.rds"))
#pred_plot_2
mse_2 <- get_mse_from_pred_plot(pred_plot_2)

```
```{r}
pred_plot_3 <- readRDS(paste0(path_to_model_folder, "/pred-plots/pred-plot-fold-3.rds"))
#pred_plot_3
mse_3 <- get_mse_from_pred_plot(pred_plot_3)

```
```{r}
pred_plot_4 <- readRDS(paste0(path_to_model_folder, "/pred-plots/pred-plot-fold-4.rds"))
#pred_plot_4
mse_4 <- get_mse_from_pred_plot(pred_plot_4)

```
```{r}
pred_plot_5 <- readRDS(paste0(path_to_model_folder, "/pred-plots/pred-plot-fold-5.rds"))
#pred_plot_5
mse_5 <- get_mse_from_pred_plot(pred_plot_5)

```
```{r}
pred_plot_list <- list(pred_plot_1,pred_plot_2,pred_plot_3,pred_plot_4,pred_plot_5)
#lapply(pred_plot_list, get_mse_from_pred_plot)
```
```{r pred-plot-fold-lasso-stand, fig.cap="Precipitation prediction and target values in the test set in each fold. Predictions in red and target values in black."}
pred_plot_1 + pred_plot_2 + pred_plot_3 + pred_plot_4 +
  pred_plot_5
```
The predictions look similar to those without standardization,
with differences for example in fold 2. With standardization the predictions 
are lower than the target in the later months. But the peak in fold 4 is better fit
by this model than by the original lasso.

## Inspect predictions from best CV-lambda

For consistency we also choose the minimizing lambda here.

<!-- ```{r} -->
<!-- # fit model with best lambda -->
<!-- #which.min(b) -->
<!-- lambdas <- readRDS("../results/CV-lasso/cv-lasso-og-data-16-06-22/lambda-vec.rds") -->
<!-- ``` -->

<!-- ```{r} -->
<!-- minmin <- which.min(apply(err_mat, 1, mean)) -->
<!-- #minmin <- l1se -->
<!-- best_lambda <- rev(lambdas)[minmin] -->
<!-- #dim(features) -->
<!-- m_full <- glmnet(features[1:370,], target[1:370], lambda = best_lambda, standardize = FALSE) -->
<!-- preds_full <- predict(m_full, newx = features[371:432,], -->
<!--                       lambda = best_lambda) -->
<!-- df <- data.frame(preds = preds_full, target = target[371:432]) -->
<!-- #dim(preds_full) -->
<!-- ggplot() + geom_line(data = df, mapping = aes(x=1:62, y=s0, colour = "blue")) + -->
<!--   geom_line(data = df, mapping= aes(x=1:62, y=target)) -->

<!-- ``` -->
<!-- ```{r} -->
<!-- sqrt(mean((df$s0 - df$target)^2)) -->
<!-- ``` -->
<!-- ```{r} -->
<!-- mean((df$s0 - df$target)^2) -->
<!-- ``` -->

```{r}
a <- readRDS(paste0(path_to_model_folder, "/pred-plots/fitted-preds-plot-full.rds"))
a
```



```{r pred-plot-full-lasso-stand, fi.cap="Precipitation prediction and target values in the validation set. Predictions in red and target values in black. The model was fitted on the full CV data with the lambda value that minimised the average MSE"}
full_preds <- readRDS(paste0(path_to_model_folder, "/pred-plots/pred-plot-full.rds"))
full_preds
mse_full <- get_mse_from_pred_plot(full_preds)
mse_full
```

```{r}
full_coef <- readRDS(paste0(path_to_model_folder, "/coef-plots/coef-plot-full.rds"))
full_coef
```


Over the more than 5 years of validation data the model predicts the
seasonal pattern of the precipitation time series quite well,
but constantly fails to predict the higher values of precipitation.
The MSE is `mse_full` and the RSME `sqrt(mse_full)``.

## Summary

By standardizing we improve the MSE of the final model slightly but
still under estimate the larger precipitation values in the evaluation set.

<!-- Also although predicting seems to bring useful results, the choice of the LASSO coefficients seems somewhat arbitrary or at least not -->
<!-- interpret able in a straightforward way. -->


<!-- ```{r} -->
<!-- m_full2 <- glmnet(features[1:370,], target[1:370], lambda = rev(lambdas), standardize = FALSE) -->
<!-- preds_full2 <- predict.glmnet(m_full2, newx = features[371:432,], s = rev(lambdas)) -->
<!-- v <- apply(preds_full2,2,function(x) mean((x-target[371:432])^2)) -->
<!-- v <- unname(v) -->
<!-- best_l <- which.min(v) -->
<!-- df2 <- data.frame(preds = preds_full2[,best_l], target = target[371:432]) -->
<!-- #dim(preds_full) -->
<!-- ggplot() + geom_line(data = df2, mapping = aes(x=1:62, y=preds, colour = "blue")) + -->
<!--   geom_line(data = df, mapping= aes(x=1:62, y=target)) -->
<!-- ``` -->
<!-- ```{r} -->
<!-- sqrt(mean((df2$preds - df$target)^2)) -->
<!-- ``` -->

```{r}
# save full model after fitting!
df <- ggplot_build(full_preds)
d <- df$data
a <- d[[1]]$y # fitted value
b <- d[[2]]$y # 
res <- a-b # residuals

plot(a,res)
qqnorm(res)
qqline(res)
plot(density(res))
```

```{r}
acf(res)
Box.test(res, lag=20, type="Ljung-Box")
```
```{r}
plot.ts(res)
```
```{r}
pacf(res)
```


```{r}
plot(hist(res))
```


