# Lasso on cluster 5

```{r, echo = FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE, echo = FALSE)
```


```{r}
library(patchwork)
library(ggpubr)
library(raster)
library(glmnet)
library(Hmisc)
source("../code/R/helper-functions.R")
```

```{r}
path_to_model_folder <- "../results/CV-lasso/cluster-cv-lasso-og2/cluster-5/"
```


## Cluster 5

These are the cluster results for cluster 5

## Error plots


```{r err-bar-plot-lasso-og-cl5, fig.cap="Mean squared error of the 5-fold blocked cross validation for cluster 5. On the x-axis the lambda values on the log scale. The points in the middle represent the average MSE for the respective lambda, the errorbars give the MSE +/- one standard deviation. The dotted line shows the lambda for which minimum MSE was obtained."}
err_bars_plot <- readRDS(paste0(path_to_model_folder, "/err-mat-plots/err-bars-plot.rds"))
err_bars_plot
```


```{r err-fold-lasso-og-cl5, fig.cap="MSE of the CV for the different lambda values on the a log scale. The red dotted line shows the lambda for which minimum MSE was obtained."}
p1 <- readRDS(paste0(path_to_model_folder, "err-mat-plots/err-plot-fold-1.rds"))
p2 <- readRDS(paste0(path_to_model_folder, "err-mat-plots/err-plot-fold-2.rds"))
p3 <- readRDS(paste0(path_to_model_folder, "err-mat-plots/err-plot-fold-3.rds"))
p4 <- readRDS(paste0(path_to_model_folder, "err-mat-plots/err-plot-fold-4.rds"))
p5 <- readRDS(paste0(path_to_model_folder, "err-mat-plots/err-plot-fold-5.rds"))

p1 + p2 + p3 + p4 + p5
```

```{r}
err_mat_list <- list(p1,p2,p3,p4,p5)
err_mat_resc_list <- replot_err_mat(err_mat_list)
```

```{r err-fold-lasso-og-replot-cl5, fig.cap="MSE of the CV for the different lambda values on the a log scale computed for cluster 5. The red dotted line shows the lambda for which minimum MSE was obtained. See @\ref(fig:err-folg-lasso-og), but this time the y-axis has the same range for all plots."}
err_mat_resc_list[[1]] + err_mat_resc_list[[2]] + err_mat_resc_list[[3]] + err_mat_resc_list[[4]] + err_mat_resc_list[[5]]
```

Below we can see the minimum MSE for each fold.
```{r}
#apply(err_mat, 2, min)
```
And which lambda in the lambda vector resulted in the lowest prediction error on the folds´ test set.

```{r}
#apply(err_mat, 2, function(x) which.min(x))
```



## Coefficient plots

```{r coef-plot-lasso-og-cl5, fig.cap="Coefficient map plot for the different folds. Longitude and Latitude on the x and y-axis respectively. Positive values are coloured in blue, negative values in red."}
pc1 <- readRDS(paste0(path_to_model_folder, "coef-plots/coef-plot-fold-1.rds"))
pc2 <- readRDS(paste0(path_to_model_folder, "coef-plots/coef-plot-fold-2.rds"))
pc3 <- readRDS(paste0(path_to_model_folder, "coef-plots/coef-plot-fold-3.rds"))
pc4 <- readRDS(paste0(path_to_model_folder, "coef-plots/coef-plot-fold-4.rds"))
pc5 <- readRDS(paste0(path_to_model_folder, "coef-plots/coef-plot-fold-5.rds"))



l <- list(pc1,pc2,pc3,pc4,pc5)
ggarrange(plotlist = l, ncol = 3, nrow = 2,
         common.legend = TRUE)
```


## Inspect predictions from each fold

Following we inspect the folds precipitation time series and
the predictions made by the model.


```{r}
pred_plot_1 <- readRDS(paste0(path_to_model_folder, "/pred-plots/pred-plot-fold-1.rds"))
pred_plot_1
```
```{r}
pred_plot_2 <- readRDS(paste0(path_to_model_folder, "/pred-plots/pred-plot-fold-2.rds"))
pred_plot_2
```
```{r}
pred_plot_3 <- readRDS(paste0(path_to_model_folder, "/pred-plots/pred-plot-fold-3.rds"))
pred_plot_3
```
```{r}
pred_plot_4 <- readRDS(paste0(path_to_model_folder, "/pred-plots/pred-plot-fold-4.rds"))
pred_plot_4
```
```{r}
pred_plot_5 <- readRDS(paste0(path_to_model_folder, "/pred-plots/pred-plot-fold-5.rds"))
pred_plot_5
```

## Inspect predictions from best CV-lambda

We choose now the best lambda and fit 
the model anew to predict the time series we held out
as validation data



```{r}
full_preds <- readRDS(paste0(path_to_model_folder, "/pred-plots/pred-plot-full.rds"))
full_preds
mse_full <- get_mse_from_pred_plot(full_preds)
mse_full
```
```{r}
# [1] 1616.594
# [1] "Done fitting and evaluating the full model with l_min for cluster 1"
# [1] 1250.799
# [1] "Done fitting and evaluating the full model with l_min for cluster 2"
# [1] 3091.448
# [1] "Done fitting and evaluating the full model with l_min for cluster 3"
# [1] 3505.33
# [1] "Done fitting and evaluating the full model with l_min for cluster 4"
# [1] 2599.921
# [1] "Done fitting and evaluating the full model with l_min for cluster 5"
```


## Summary
