## Without centering

### K-means and PAM gap statistics without PCA 
```{r, include = FALSE}
gap_kmeans_plot <- readRDS("../results/clustering/gap_kmeans_plot.rds")

```

```{r, include = FALSE}
gap_pam_plot <- readRDS("../results/clustering/gap_pam_plot.rds")

```

```{r}
gap_kmeans_plot + gap_pam_plot
```

As for the centered data, K-means and PAM choose the maximum number
of possible clusters as optimal.


### Scree plot
```{r}
scree_plot_pca_uncentered <- readRDS("../results/scree_plot_pca_uncentered.rds")
scree_plot_pca_uncentered
```

We now apply a PCA on the uncentered data.
The first principal component now explains a lot more variance.
See https://eigenvector.com/wp-content/uploads/2020/06/EffectofCenteringonPCA.pdf

We proceed as for the centered data but use 2 and 3 principal components respectively

### Kmean and PAM gap statistics after applying PCA


```{r}
summary_plot_pca_uncentered <- readRDS("../results/clustering/summary_plot_pca_uncentered.rds")
summary_plot_pca_uncentered
```

The graphic shows the gap statistics resulting from K-means and PAM for 2 and 3 principal components respectively. For 2 principal components
K-means and PAM both choose 3 as the optimal number of clusters, while
for 3 principal components hey choose 12.
The gap statistic "spike" for 2 principal components is more distinct.



