# Results LASSO og center

```{r, echo = FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE, echo = FALSE)
```


```{r}
library(patchwork)
library(ggpubr)
library(raster)
library(glmnet)
library(Hmisc)
source("../code/R/helper-functions.R")
```

```{r}
# inspect error plots
# inspect coef plots
# inspect prediction plots
# length(ids$train$Training060)
# length(ids$test$Testing060)
```

## LASSO model

We fit a LASSO model on the precipitation and SST data.
The precipitation target is the monthly mean precipitation
in the Central Amazon Basin, the SST data are monthly temperatures
over the globe.
We use a 5-fold CV approach to find an optimal lambda.
Each fold consists of 5 consecutive years of training data
followed by 2 years of test data.
In each fold we fit a LASSO model on a set of predetermined lambda 
values and choose the lambda that minimizes the MSE on the test set 
in that fold.
After determining the best lambda in each fold we choose the
lambda that minimizes the MSE over all folds and refit the model
to the complete training data. Afterwards we evaluate the fitted model on a separate validation set with 5 years length which was not included
in the training phase.

## Error plots

```{r}
err_mat <- readRDS("../results/CV-lasso/cv-lasso-og-center-11-06-22/err-mat.rds")
#err_mat <- sqrt(err_mat)
err_mat <- err_mat[1:60,]
```

```{r}
lambda_id <- c(1:60)
mse <- apply(err_mat, 1, mean)
b <- mse
bplus <- apply(err_mat, 1, max)
bmin <- apply(err_mat, 1, min)
errbar(lambda_id, mse, b+bplus, b-bmin)
```

```{r}
# which.min(b)
# bs <- b[40:100]
# bbpluss <- (b+bplus)[40:100]
# bbmins <- (b-bmin)[40:100]
# (bs > bbpluss) | (bs < bbmins)
bmin <- which.min(b)
tv <- (b+bplus)[bmin]
tz <- b+bplus
tz <- tz[(bmin+1):100]
w <- tv < tz
#bmin is 40
#w ist bei 45 true
#l1se <- 45

```



```{r}
p1 <- readRDS("../results/CV-lasso/cv-lasso-og-center-11-06-22/err-mat-plots/err-plot-fold-1.rds")
p2 <- readRDS("../results/CV-lasso/cv-lasso-og-center-11-06-22/err-mat-plots/err-plot-fold-2.rds")
p3 <- readRDS("../results/CV-lasso/cv-lasso-og-center-11-06-22/err-mat-plots/err-plot-fold-3.rds")
p4 <- readRDS("../results/CV-lasso/cv-lasso-og-center-11-06-22/err-mat-plots/err-plot-fold-4.rds")
p5 <- readRDS("../results/CV-lasso/cv-lasso-og-center-11-06-22/err-mat-plots/err-plot-fold-5.rds")
p1 + p2 + p3 + p4 + p5
```

```{r}
apply(err_mat, 2, min)
```
And which lambda in the lambda vector resulted in the lowest prediction error on the folds´ test set.

```{r}
apply(err_mat, 2, function(x) which.min(x))
```



## Coefficient plots

```{r}
pc1 <- readRDS("../results/CV-lasso/cv-lasso-og-center-11-06-22/coef-plots/coef-plot-fold-1.rds")
pc2 <- readRDS("../results/CV-lasso/cv-lasso-og-center-11-06-22/coef-plots/coef-plot-fold-2.rds")
pc3 <- readRDS("../results/CV-lasso/cv-lasso-og-center-11-06-22/coef-plots/coef-plot-fold-3.rds")
pc4 <- readRDS("../results/CV-lasso/cv-lasso-og-center-11-06-22/coef-plots/coef-plot-fold-4.rds")
pc5 <- readRDS("../results/CV-lasso/cv-lasso-og-center-11-06-22/coef-plots/coef-plot-fold-5.rds")

l <- list(pc1,pc2,pc3,pc4,pc5)
ggarrange(plotlist = l, ncol = 3, nrow = 2,
          common.legend = TRUE)
```



```{r}
# q <- ggplot_build(pc1)
# q$data[[1]]$linetype <- 21
# pc1$layers[[2]]$aes_params$shape <- 21
# pc1$data
# pc1$layers[[2]]
# q <- ggplot_gtable(q)
# plot(q)
# 
# q <- ggplot_build(pc1)
```



## Inspect predictions from each fold

Following we inspect the folds precipitation time series and
the predictions made by the model.

```{r}
# load data
target_path <- "../data/interim/drought/chirps_setreftime_aggregated.rds"
features_path <- "../data/interim/sst/ersst_setreftime.nc"
target <- load_data(target_path)
target <- apply(getValues(target), 2, mean)
features <- load_data(features_path, "sst")
features <- t(features)
features <- prepare_sst(features)
features <- scale(features, center = TRUE, scale = FALSE)
```

```{r}
#plot(ts(target))
```


```{r}
# get indices
lambdas <- readRDS("../results/CV-lasso/cv-lasso-og-center-11-06-22/lambda-vec.rds")
ids <- readRDS("../results/CV-lasso/cv-lasso-og-center-11-06-22/index-list.rds")
```


```{r}
f1_train <- ids$train$Training060
f1_test <- ids$test$Testing060
x1_train <- features[f1_train,]
y1_train <- target[f1_train]
x1_test <- features[f1_test,]
y1_test <- target[f1_test]
min1 <- which.min(err_mat[,1])
m1 <- glmnet(x1_train, y1_train, lambda=rev(lambdas)[min1], standardize = FALSE)
preds1 <- predict(object = m1, newx = x1_test)
df <- data.frame(preds = preds1, target = y1_test, check.names=TRUE)
ggplot() + geom_line(data = df, mapping = aes(x=1:14, y=s0, colour = "blue")) +
  geom_line(data = df, mapping= aes(x=1:14, y=target))
```

```{r}
f2_train <- ids$train$Training134
f2_test <- ids$test$Testing134
x2_train <- features[f2_train,]
y2_train <- target[f2_train]
x2_test <- features[f2_test,]
y2_test <- target[f2_test]
min2 <- which.min(err_mat[,2])
m2 <- glmnet(x2_train, y2_train, lambda=rev(lambdas)[min2], standardize = FALSE)
preds2 <- predict(m2, newx = x2_test)
df <- data.frame(preds = preds2, target = y2_test)
ggplot() + geom_line(data = df, mapping = aes(x=1:14, y=s0, colour = "blue")) +
  geom_line(data = df, mapping= aes(x=1:14, y=target))
```
```{r}
f3_train <- ids$train$Training208
f3_test <- ids$test$Testing208
x3_train <- features[f3_train,]
y3_train <- target[f3_train]
x3_test <- features[f3_test,]
y3_test <- target[f3_test]
min3 <- which.min(err_mat[,3])
m3 <- glmnet(x3_train, y3_train, lambda=rev(lambdas)[min3], standardize = FALSE)
preds3 <- predict(m3, newx = x3_test)
df <- data.frame(preds = preds3, target = y3_test)
ggplot() + geom_line(data = df, mapping = aes(x=1:14, y=s0, colour = "blue")) +
  geom_line(data = df, mapping= aes(x=1:14, y=target))
```
```{r}
f4_train <- ids$train$Training282
f4_test <- ids$test$Testing282
x4_train <- features[f4_train,]
y4_train <- target[f4_train]
x4_test <- features[f4_test,]
y4_test <- target[f4_test]
min4 <- which.min(err_mat[,4])
m4 <- glmnet(x4_train, y4_train, lambda=rev(lambdas)[min4], standardize = FALSE)
preds4 <- predict(m4, newx = x4_test)
df <- data.frame(preds = preds4, target = y4_test)
ggplot() + geom_line(data = df, mapping = aes(x=1:14, y=s0, colour = "blue")) +
  geom_line(data = df, mapping= aes(x=1:14, y=target))
```

```{r}
f5_train <- ids$train$Training356
f5_test <- ids$test$Testing356
x5_train <- features[f5_train,]
y5_train <- target[f5_train]
x5_test <- features[f5_test,]
y5_test <- target[f5_test]
min5 <- which.min(err_mat[,5])
m5 <- glmnet(x5_train, y5_train, lambda=rev(lambdas)[min5], standardize = FALSE)
preds5 <- predict(m5, newx = x5_test)
df <- data.frame(preds = preds5, target = y5_test)
ggplot() + geom_line(data = df, mapping = aes(x=1:14, y=s0, colour = "blue")) +
  geom_line(data = df, mapping= aes(x=1:14, y=target))
```

In general the model fits the data sufficiently to predict the general
form of the time series but misses some modes and is off in the larger
values in fold 2.

## Inspect predictions from best CV-lambda

We choose now the best lambda with the 1SE rule and fit 
the model anew to predict the time series we held out
as validation data

```{r}
# fit model with best lambda
#which.min(b)
lambdas <- readRDS("../results/CV-lasso/cv-lasso-og-center-11-06-22/lambda-vec.rds")
```

```{r}
minmin <- which.min(apply(err_mat, 1, mean))
#minmin <- l1se
best_lambda <- rev(lambdas)[minmin]
#dim(features)
m_full <- glmnet(features[1:370,], target[1:370], lambda = best_lambda, standardize = FALSE)
preds_full <- predict(m_full, newx = features[371:432,],
                      lambda = best_lambda)
df <- data.frame(preds = preds_full, target = target[371:432])
#dim(preds_full)
ggplot() + geom_line(data = df, mapping = aes(x=1:62, y=s0, colour = "blue")) +
  geom_line(data = df, mapping= aes(x=1:62, y=target))

```
```{r}
sqrt(mean((df$s0 - df$target)^2))
```
```{r}
mean((df$s0 - df$target)^2)
```

```{r}
m_full2 <- glmnet(features[1:370,], target[1:370], lambda = rev(lambdas), standardize = FALSE)
preds_full2 <- predict.glmnet(m_full2, newx = features[371:432,], s = rev(lambdas))
v <- apply(preds_full2,2,function(x) mean((x-target[371:432])^2))
v <- unname(v)
best_l <- which.min(v)
df2 <- data.frame(preds = preds_full2[,best_l], target = target[371:432])
#dim(preds_full)
ggplot() + geom_line(data = df2, mapping = aes(x=1:62, y=preds, colour = "blue")) +
  geom_line(data = df, mapping= aes(x=1:62, y=target))


```

