# glyph plots
```{r}
#pathtoplots = ptp
ptp <- "../../results/"
make_path <- function(plot_name, path_to_plots = path_to_plots) {
  a <- paste0(path_to_plots, plot_name)
  return(a)
}
```


```{r}
seasonal_plot <- readRDS(paste0(ptp, "seasonal_plot.rds"))
seasonal_plot
```

The above figure is a glyph-map of seasonal precipitation
patterns (averages for each month) in the Central Amazon Basin.
The grey reference lines show the mid-range for easier comparison
of the patterns.
We see differences in the seasonal patterns across the map.
In the upper left for example, the seasonal patterns stay
above mid-range while on the bottom-left they have values
clearly towards the low end of the range. 
Also some areas have multimodal patterns.
The patterns differ in range and month of maximum and minimun precipitation.



```{r}
smoothed_plot <- readRDS(paste0(ptp, "smoothed_plot.rds"))
smoothed_plot
```

This plot shows the smoothed de-seasonalized monthly precipitation, after global scaling. The same position
within each cell corresponds to the same value in all locations.
Some areas have almost a linear course, increasing, decreasing
or constant. Others show a more "wiggly" courses.
As overall pattern we can see that the forms of the patterns
have a spatial connection, patterns are close to similar patterns,
at the same latitude.
Also regarding latitude the closer to the equator the less precipitation.

```{r}
smoothed_scaled_colour_plot <- readRDS(paste0(ptp, "smoothed_scaled_colour_plot.rds"))
smoothed_scaled_colour_plot
```
Now we inspect the glyph-map with deseasonalised locally scaled values. This form of scaling emphasizes the individual shapes.
Because of the applied scaling, big patterns may be just be tiny effects. Therefore colors are added according to range.
Areas with lighter colour have larger ranges than darker areas.
The areas with steep linear increases and decreases have smaller ranger than or example the areas below -2.5 latitude.


