# Fused Lasso CV

```{r, echo = FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE, echo = FALSE)
```


```{r}
library(patchwork)
library(ggpubr)
library(raster)
library(glmnet)
library(Hmisc)
library(genlasso)
source("../code/R/helper-functions.R")
```

```{r}
# inspect error plots
# inspect coef plots
# inspect prediction plots
# length(ids$train$Training060)
# length(ids$test$Testing060)
```

```{r}
f1 <- readRDS("../results/CV-lasso/fused-cv-test-100-steps/fold-models/model-fold-1.rds")
f2 <- readRDS("../results/CV-lasso/fused-cv-test-100-steps/fold-models/model-fold-2.rds")
f3 <- readRDS("../results/CV-lasso/fused-cv-test-100-steps/fold-models/model-fold-3.rds")
f4 <- readRDS("../results/CV-lasso/fused-cv-test-100-steps/fold-models/model-fold-4.rds")
f5 <- readRDS("../results/CV-lasso/fused-cv-test-100-steps/fold-models/model-fold-5.rds")
```



## Error plots

```{r}
err_mat <- readRDS("../results/CV-lasso/fused-cv-test-100-steps/err-mat.rds")
#err_mat <- sqrt(err_mat)
#err_mat <- err_mat[1:60,]
```

```{r}
lambda_id <- c(1:100)
mse <- apply(err_mat, 1, mean)
b <- mse
bplus <- apply(err_mat, 1, max)
bmin <- apply(err_mat, 1, min)
errbar(lambda_id, mse, b+bplus, b-bmin)
```

The plot shows the results from the 5 fold-CV plotted for each lambda.
Lambda values start with high regularization and then decrease.
Initially reducing regularization reduces the error
until the 43th lambda value.
As the errors get very large after the 60th lambda value and strictly
increase we only showed the first 60 values for better visibility
The highest regularization is chosen so that all coefficients are zero.



```{r}
# which.min(b)
# bs <- b[40:100]
# bbpluss <- (b+bplus)[40:100]
# bbmins <- (b-bmin)[40:100]
# (bs > bbpluss) | (bs < bbmins)
bmin <- which.min(b)
tv <- (b+bplus)[bmin]
tz <- b+bplus
tz <- tz[(bmin+1):100]
w <- tv < tz
#bmin is 40
#w ist bei 45 true
l1se <- 45

```



```{r}
# p1 <- readRDS("../results/CV-lasso/fused-cv-test-100-steps/err-mat-plots/err-plot-fold-1.rds")
# p2 <- readRDS("../results/CV-lasso/fused-cv-test-100-steps/err-mat-plots/err-plot-fold-2.rds")
# p3 <- readRDS("../results/CV-lasso/fused-cv-test-100-steps/err-mat-plots/err-plot-fold-3.rds")
# p4 <- readRDS("../results/CV-lasso/fused-cv-test-100-steps/err-mat-plots/err-plot-fold-4.rds")
# p5 <- readRDS("../results/CV-lasso/fused-cv-test-100-steps/err-mat-plots/err-plot-fold-5.rds")
# p1 + p2 + p3 + p4 + p5
plot(err_mat[,1])
```


Below we can see the minimum MSE for each fold.
```{r}
apply(err_mat, 2, min)
```
And which lambda in the lambda vector resulted in the lowest prediction error on the folds´ test set.

```{r}
apply(err_mat, 2, function(x) which.min(x))
```



## Coefficient plots

```{r}
# pc1 <- readRDS("../results/CV-lasso/fused-cv-test-100-steps/coef-plots/coef-plot-fold-1.rds")
# pc2 <- readRDS("../results/CV-lasso/fused-cv-test-100-steps/coef-plots/coef-plot-fold-2.rds")
# pc3 <- readRDS("../results/CV-lasso/fused-cv-test-100-steps/coef-plots/coef-plot-fold-3.rds")
# pc4 <- readRDS("../results/CV-lasso/fused-cv-test-100-steps/coef-plots/coef-plot-fold-4.rds")
# pc5 <- readRDS("../results/CV-lasso/fused-cv-test-100-steps/coef-plots/coef-plot-fold-5.rds")
# 
# l <- list(pc1,pc2,pc3,pc4,pc5)
# ggarrange(plotlist = l, ncol = 3, nrow = 2,
#          common.legend = TRUE)

sst <- brick("../data/interim/sst/ersst_setreftime.nc", var = "sst")
ext <- extent(-180,0,-50,40)
sst <- crop(sst,ext)
coord <- coordinates(sst)
cnames <- paste(coord[,1], coord[,2])
sst <- t(as.matrix(sst))
colnames(sst) <- cnames
sst <- prepare_sst(sst)
num_coef_names <- coef_names_to_numeric(colnames(sst))
```


```{r}
co1 <- coef.genlasso(f1)$beta
#l1 <- coef.genlasso(f1)$lambda[which.min(err_mat[,1])]
l1 <- which.min(err_mat[,1])
co1 <- co1[,l1]
coef_mat <- cbind(num_coef_names, co1)
plot1 <- plot_nonzero_coefficients(coef_mat)
plot1
# range(co1)
# table(round(co1,3))

# preapare coefficients from fold
# get coordiante names from sst
# cbind to get coef_matrix
# plot with plot_nonzero
# 
# plot_fused_fold <- function(fold_model) {
#   
# }


p1s <- softthresh(f1, lambda=4069.727,gamma=0.0001)
plot(p1s)
```

```{r}
co2 <- coef.genlasso(f2)$beta
#l1 <- coef.genlasso(f1)$lambda[which.min(err_mat[,1])]
l2 <- which.min(err_mat[,2])
co2 <- co2[,l2]
coef_mat <- cbind(num_coef_names, co2)
plot2 <- plot_nonzero_coefficients(coef_mat)
plot2
range(co1)
table(round(co1,3))
```

```{r}
co3 <- coef.genlasso(f3)$beta
#l1 <- coef.genlasso(f1)$lambda[which.min(err_mat[,1])]
l3 <- which.min(err_mat[,3])
co3 <- co3[,l3]
coef_mat <- cbind(num_coef_names, co3)
plot3 <- plot_nonzero_coefficients(coef_mat)
plot3
range(co1)
table(round(co1,3))
```

```{r}
co4 <- coef.genlasso(f4)$beta
#l1 <- coef.genlasso(f1)$lambda[which.min(err_mat[,1])]
l4 <- which.min(err_mat[,4])
co4 <- co4[,l4]
coef_mat <- cbind(num_coef_names, co4)
plot4 <- plot_nonzero_coefficients(coef_mat)
plot4
range(co1)
table(round(co1,3))
```

```{r}
co5 <- coef.genlasso(f5)$beta
#l1 <- coef.genlasso(f1)$lambda[which.min(err_mat[,1])]
l5 <- which.min(err_mat[,5])
co5 <- co5[,l5]
coef_mat <- cbind(num_coef_names, co5)
plot5 <- plot_nonzero_coefficients(coef_mat)
plot5
range(co1)
table(round(co1,3))
```


```{r}
# q <- ggplot_build(pc1)
# q$data[[1]]$linetype <- 21
# pc1$layers[[2]]$aes_params$shape <- 21
# pc1$data
# pc1$layers[[2]]
# q <- ggplot_gtable(q)
# plot(q)
# 
# q <- ggplot_build(pc1)
```



## Inspect predictions from each fold

```{r}
# load model
# evaluate predictions on the test data
target_path <- "../data/interim/drought/chirps_setreftime_aggregated.rds"
features_path <- "../data/interim/sst/ersst_setreftime.nc"
target <- load_data(target_path)
target <- apply(getValues(target), 2, mean)
sst <- brick("../data/interim/sst/ersst_setreftime.nc", var = "sst")
ext <- extent(-180,0,-50,40)
sst <- crop(sst,ext)
features <- as.matrix(sst)
features <- t(features)
features <- prepare_sst(features)
dim(features)
ids <- readRDS("../results/CV-lasso/cv-lasso-og-data-31-05-22/index-list.rds")
```

```{r}
f1 <- readRDS("../results/CV-lasso/fused-cv-test-100-steps/fold-models/model-fold-1.rds")
preds1 <- predict.genlasso(f1, Xnew=features[ids$test$Testing060,])
dim(preds1$fit)
min1 <- which.min(err_mat[,1])
p1 <- preds1$fit[,1]
df <- data.frame(preds = p1, target = target[ids$test$Testing060])
ggplot() + geom_line(data = df, mapping = aes(x=1:14, y=preds, colour = "blue")) +
  geom_line(data = df, mapping= aes(x=1:14, y=target))
```
```{r}
f2 <- readRDS("../results/CV-lasso/fused-cv-test-100-steps/fold-models/model-fold-2.rds")
preds2 <- predict.genlasso(f2, Xnew=features[ids$test$Testing134,])
dim(preds2$fit)
min2 <- which.min(err_mat[,2])
p2 <- preds2$fit[,2]
df <- data.frame(preds = p2, target = target[ids$test$Testing134])
ggplot() + geom_line(data = df, mapping = aes(x=1:14, y=preds, colour = "blue")) +
  geom_line(data = df, mapping= aes(x=1:14, y=target))
```

```{r}
f3 <- readRDS("../results/CV-lasso/fused-cv-test-100-steps/fold-models/model-fold-3.rds")
preds3 <- predict.genlasso(f3, Xnew=features[ids$test$Testing208,])
dim(preds3$fit)
min3 <- which.min(err_mat[,3])
p3 <- preds3$fit[,3]
df <- data.frame(preds = p3, target = target[ids$test$Testing208])
ggplot() + geom_line(data = df, mapping = aes(x=1:14, y=preds, colour = "blue")) +
  geom_line(data = df, mapping= aes(x=1:14, y=target))

```
```{r}
f4 <- readRDS("../results/CV-lasso/fused-cv-test-100-steps/fold-models/model-fold-4.rds")
preds4 <- predict.genlasso(f4, Xnew=features[ids$test$Testing282,])
min4 <- which.min(err_mat[,4])
p4 <- preds4$fit[,4]
df <- data.frame(preds = p4, target = target[ids$test$Testing282])
ggplot() + geom_line(data = df, mapping = aes(x=1:14, y=preds, colour = "blue")) +
  geom_line(data = df, mapping= aes(x=1:14, y=target))

```

```{r}
f5 <- readRDS("../results/CV-lasso/fused-cv-test-100-steps/fold-models/model-fold-5.rds")
preds5 <- predict.genlasso(f5, Xnew=features[ids$test$Testing356,])
min5 <- which.min(err_mat[,5])
p5 <- preds5$fit[,5]
df <- data.frame(preds = p5, target = target[ids$test$Testing356])
ggplot() + geom_line(data = df, mapping = aes(x=1:14, y=preds, colour = "blue")) +
  geom_line(data = df, mapping= aes(x=1:14, y=target))
```


```{r}
# fit model with best lambda
#which.min(b)
#lambdas <- readRDS("../results/CV-lasso/fused-cv-test-100-steps/lambda-vec.rds")
```

```{r}
# full model prediction for fused lasso cv in other script
```
```{r}
#sqrt(mean((df$s0 - df$target)^2))
```
```{r}
#mean((df$s0 - df$target)^2)
```
Over the more than 5 years of validation data the model predicts the
seasonal pattern of the precipitation time series quite well,
but constantly fails to predict the higher values of precipitation.
The MSE is 1215.74 and the RSME 34.87.

## Summary
We fitted a LASSO model for predicting the mean precipitation
in the Central Amazon Basin and used a 5-fold blocked Cross Validation
approach to find the optimal level of regularization.
After training the model we evaluated its performance on a separate validation set that was not used in the training process.
The model shows predicting capabilities but misses out on higher values of the precipitation target.
Also although predicting seems to bring useful results, the choice of the LASSO coefficients seems somewhat arbitrary or at least not
interpret able in a straightforward way.


