# Glyph plots
```{r, echo=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(warning = FALSE)
```


```{r}
#pathtoplots = ptp
library(ggplot2)
library(ggtext)
ptp <- "../results/"
```


```{r, fig.cap="Glyph map of seasonal precipitaton pattern. Each location is presented by a time series. The time series are seperated by boxes. The gray reference lines inside the boxes show the mid-range for easier comparison."}
seasonal_plot <- readRDS(paste0(ptp, "seasonal_plot.rds"))
seasonal_plot + ggtitle("Glyph map of seasonal precipitation patterns (averages for each month)") +
  theme(plot.title = element_textbox_simple())
```

The above figure is a glyph-map of seasonal precipitation
patterns (averages for each month) in the Central Amazon Basin.
The gray reference lines show the mid-range for easier comparison
of the patterns.
We see differences in the seasonal patterns across the map.
In the upper left for example, the seasonal patterns stay
above mid-range while on the bottom-left they have values
clearly towards the low end of the range. 
Also some areas have multimodal patterns.
The patterns differ in range and month of maximum and minimum precipitation.



```{r, fig.cap="Glyph map of de-seasonalised and smoothed precipitation. Each location is presented by a time series. The time series are seperated by boxes. The gray reference lines inside the boxes show the mid-range for easier comparison. The time series are scaled globally, same positions inside the cells correspond to the same values in all locations."}
smoothed_plot <- readRDS(paste0(ptp, "smoothed_plot.rds"))
smoothed_plot + ggtitle("Glyph map of smoothed de-seasonalised monthly precipitation") +
  theme(plot.title = element_textbox_simple())
```

This plot shows the smoothed de-seasonalized monthly precipitation, after global scaling. The same position
within each cell corresponds to the same value in all locations.
Some areas have almost a linear course, increasing, decreasing
or constant. Others show a more "wiggly" courses.
As overall pattern we can see that the forms of the patterns
have a spatial connection, patterns are close to similar patterns,
at the same latitude.
Also regarding latitude the closer to the equator the less precipitation.

```{r, fig.cap="Glyph map of de-seasonalised and smoothed precipitation. The time series are scaled locally, ranges are not the same in all cells. The different ranges are given in color shades, where lighter shading indicates a larger range and darker shades smaller ranges."}
smoothed_scaled_colour_plot <- readRDS(paste0(ptp, "smoothed_scaled_colour_plot.rds"))
smoothed_scaled_colour_plot + ggtitle("Glyph map of smoothed de-seasonalised monthly precipitation, locally scaled") +
  theme(plot.title = element_textbox_simple()) +
  theme(legend.position = "bottom") 
```

Now we inspect the glyph-map with de-seasonalized locally scaled values. This form of scaling emphasizes the individual shapes.
Because of the applied scaling, big patterns may be just be tiny effects. Therefore colors are added according to range.
Areas with lighter color have larger ranges than darker areas.
The areas with steep linear increases and decreases have smaller ranges than or example the areas below -2.5 latitude in the left.


