# Lasso on original data

```{r, echo = FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE, echo = FALSE)
```


```{r}
library(patchwork)
library(ggpubr)
library(raster)
library(glmnet)
library(Hmisc)
source("../code/R/helper-functions.R")
```

```{r}
# inspect error plots
# inspect coef plots
# inspect prediction plots
# length(ids$train$Training060)
# length(ids$test$Testing060)
```

## LASSO model

We fit a LASSO model on the precipitation and SST data.
The precipitation target is the monthly mean precipitation
in the Central Amazon Basin, the SST data are monthly temperatures
over the globe.
We use a 5-fold CV approach to find an optimal lambda.
Each fold consists of 5 consecutive years of training data
followed by 2 years of test data.
In each fold we fit a LASSO model on a set of predetermined lambda 
values and choose the lambda that minimizes the MSE on the test set 
in that fold.
After determining the best lambda in each fold we choose the
lambda that minimizes the MSE over all folds and refit the model
to the complete training data. Afterwards we evaluate the fitted model on a separate validation set with 5 years length which was not included
in the training phase.

## Error plots

```{r}
err_mat <- readRDS("../results/CV-lasso/cv-lasso-og-data-31-05-22/err-mat.rds")
#err_mat <- sqrt(err_mat)
err_mat <- err_mat[1:60,]
```

```{r}
lambda_id <- c(1:60)
mse <- apply(err_mat, 1, mean)
b <- mse
bplus <- apply(err_mat, 1, max)
bmin <- apply(err_mat, 1, min)
errbar(lambda_id, mse, b+bplus, b-bmin)
```

The plot shows the results from the 5 fold-CV plotted for each lambda.
Lambda values start with high regularization and then decrease.
Initially reducing regularization reduces the error
until the 43th lambda value.
As the errors get very large after the 60th lambda value and strictly
increase we only showed the first 60 values for better visibility
The highest regularization is chosen so that all coefficients are zero.



```{r}
# which.min(b)
# bs <- b[40:100]
# bbpluss <- (b+bplus)[40:100]
# bbmins <- (b-bmin)[40:100]
# (bs > bbpluss) | (bs < bbmins)
bmin <- which.min(b)
tv <- (b+bplus)[bmin]
tz <- b+bplus
tz <- tz[(bmin+1):100]
w <- tv < tz
#bmin is 40
#w ist bei 45 true
l1se <- 45

```



```{r}
p1 <- readRDS("../results/CV-lasso/cv-lasso-og-data-31-05-22/err-mat-plots/err-plot-fold-1.rds")
p2 <- readRDS("../results/CV-lasso/cv-lasso-og-data-31-05-22/err-mat-plots/err-plot-fold-2.rds")
p3 <- readRDS("../results/CV-lasso/cv-lasso-og-data-31-05-22/err-mat-plots/err-plot-fold-3.rds")
p4 <- readRDS("../results/CV-lasso/cv-lasso-og-data-31-05-22/err-mat-plots/err-plot-fold-4.rds")
p5 <- readRDS("../results/CV-lasso/cv-lasso-og-data-31-05-22/err-mat-plots/err-plot-fold-5.rds")
p1 + p2 + p3 + p4 + p5
```

In the above plot we can see the error curves for each fold.
Folds 1,2 and 3 in the top row, left to right and folds 4 and 5 in the bottom row.
We can see that initially that with decreasing regularization the MSE decreases as well.
Up to a certain point after which the error rises again.
Fold 4 seems to be the exception where a decrease in regularization
does not increase the error dramatically as seen for example in folds 2 and 5 respectively.

Below we can see the minimum MSE for each fold.
```{r}
apply(err_mat, 2, min)
```
And which lambda in the lambda vector resulted in the lowest prediction error on the folds´ test set.

```{r}
apply(err_mat, 2, function(x) which.min(x))
```



## Coefficient plots

```{r}
pc1 <- readRDS("../results/CV-lasso/cv-lasso-og-data-31-05-22/coef-plots/coef-plot-fold-1.rds")
pc2 <- readRDS("../results/CV-lasso/cv-lasso-og-data-31-05-22/coef-plots/coef-plot-fold-2.rds")
pc3 <- readRDS("../results/CV-lasso/cv-lasso-og-data-31-05-22/coef-plots/coef-plot-fold-3.rds")
pc4 <- readRDS("../results/CV-lasso/cv-lasso-og-data-31-05-22/coef-plots/coef-plot-fold-4.rds")
pc5 <- readRDS("../results/CV-lasso/cv-lasso-og-data-31-05-22/coef-plots/coef-plot-fold-5.rds")

l <- list(pc1,pc2,pc3,pc4,pc5)
ggarrange(plotlist = l, ncol = 3, nrow = 2,
         common.legend = TRUE)
```

The plots displays the nonzero coefficients in each fold computed
for the lambda that minimizes the MSE on the test set in the respective fold.
For all folds we see large negative coefficients in the arctic. 
The LASSO chooses among correlated variables only one and discards the others,
which can be seen here since the variables chosen are scattered across the map and can but don't have to be close to each other.


```{r}
# q <- ggplot_build(pc1)
# q$data[[1]]$linetype <- 21
# pc1$layers[[2]]$aes_params$shape <- 21
# pc1$data
# pc1$layers[[2]]
# q <- ggplot_gtable(q)
# plot(q)
# 
# q <- ggplot_build(pc1)
```



## Inspect predictions from each fold

Following we inspect the folds precipitation time series and
the predictions made by the model.

```{r}
# load data
target_path <- "../data/interim/drought/chirps_setreftime_aggregated.rds"
features_path <- "../data/interim/sst/ersst_setreftime.nc"
target <- load_data(target_path)
target <- apply(getValues(target), 2, mean)
features <- load_data(features_path, "sst")
features <- t(features)
features <- prepare_sst(features)
```

```{r}
#plot(ts(target))
```


```{r}
# get indices
lambdas <- readRDS("../results/CV-lasso/cv-lasso-og-data-31-05-22/lambda-vec.rds")
ids <- readRDS("../results/CV-lasso/cv-lasso-og-data-31-05-22/index-list.rds")
```


```{r}
f1_train <- ids$train$Training060
f1_test <- ids$test$Testing060
x1_train <- features[f1_train,]
y1_train <- target[f1_train]
x1_test <- features[f1_test,]
y1_test <- target[f1_test]
min1 <- which.min(err_mat[,1])
m1 <- glmnet(x1_train, y1_train, lambda=rev(lambdas)[min1])
preds1 <- predict(object = m1, newx = x1_test)
df <- data.frame(preds = preds1, target = y1_test, check.names=TRUE)
ggplot() + geom_line(data = df, mapping = aes(x=1:14, y=s0, colour = "blue")) +
  geom_line(data = df, mapping= aes(x=1:14, y=target))
```

```{r}
f2_train <- ids$train$Training134
f2_test <- ids$test$Testing134
x2_train <- features[f2_train,]
y2_train <- target[f2_train]
x2_test <- features[f2_test,]
y2_test <- target[f2_test]
min2 <- which.min(err_mat[,2])
m2 <- glmnet(x2_train, y2_train, lambda=rev(lambdas)[min2])
preds2 <- predict(m2, newx = x2_test)
df <- data.frame(preds = preds2, target = y2_test)
ggplot() + geom_line(data = df, mapping = aes(x=1:14, y=s0, colour = "blue")) +
  geom_line(data = df, mapping= aes(x=1:14, y=target))
```
```{r}
f3_train <- ids$train$Training208
f3_test <- ids$test$Testing208
x3_train <- features[f3_train,]
y3_train <- target[f3_train]
x3_test <- features[f3_test,]
y3_test <- target[f3_test]
min3 <- which.min(err_mat[,3])
m3 <- glmnet(x3_train, y3_train, lambda=rev(lambdas)[min3])
preds3 <- predict(m3, newx = x3_test)
df <- data.frame(preds = preds3, target = y3_test)
ggplot() + geom_line(data = df, mapping = aes(x=1:14, y=s0, colour = "blue")) +
  geom_line(data = df, mapping= aes(x=1:14, y=target))
```
```{r}
f4_train <- ids$train$Training282
f4_test <- ids$test$Testing282
x4_train <- features[f4_train,]
y4_train <- target[f4_train]
x4_test <- features[f4_test,]
y4_test <- target[f4_test]
min4 <- which.min(err_mat[,4])
m4 <- glmnet(x4_train, y4_train, lambda=rev(lambdas)[min4])
preds4 <- predict(m4, newx = x4_test)
df <- data.frame(preds = preds4, target = y4_test)
ggplot() + geom_line(data = df, mapping = aes(x=1:14, y=s0, colour = "blue")) +
  geom_line(data = df, mapping= aes(x=1:14, y=target))
```

```{r}
f5_train <- ids$train$Training356
f5_test <- ids$test$Testing356
x5_train <- features[f5_train,]
y5_train <- target[f5_train]
x5_test <- features[f5_test,]
y5_test <- target[f5_test]
min5 <- which.min(err_mat[,5])
m5 <- glmnet(x5_train, y5_train, lambda=rev(lambdas)[min5])
preds5 <- predict(m5, newx = x5_test)
df <- data.frame(preds = preds5, target = y5_test)
ggplot() + geom_line(data = df, mapping = aes(x=1:14, y=s0, colour = "blue")) +
  geom_line(data = df, mapping= aes(x=1:14, y=target))
```

In general the model fits the data sufficiently to predict the general
form of the time series but misses some modes and is off in the larger
values in fold 2.

## Inspect predictions from best CV-lambda

We choose now the best lambda with the 1SE rule and fit 
the model anew to predict the time series we held out
as validation data

```{r}
# fit model with best lambda
#which.min(b)
lambdas <- readRDS("../results/CV-lasso/cv-lasso-og-data-31-05-22/lambda-vec.rds")
```

```{r}
#minmin <- which.min(apply(err_mat, 1, mean))
minmin <- l1se
best_lambda <- rev(lambdas)[minmin]
#dim(features)
m_full <- glmnet(features[1:370,], target[1:370], lambda = best_lambda)
preds_full <- predict(m_full, newx = features[371:432,],
                      lambda = best_lambda)
df <- data.frame(preds = preds_full, target = target[371:432])
#dim(preds_full)
ggplot() + geom_line(data = df, mapping = aes(x=1:62, y=s0, colour = "blue")) +
  geom_line(data = df, mapping= aes(x=1:62, y=target))

```
```{r}
#sqrt(mean((df$s0 - df$target)^2))
```
```{r}
#mean((df$s0 - df$target)^2)
```
Over the more than 5 years of validation data the model predicts the
seasonal pattern of the precipitation time series quite well,
but constantly fails to predict the higher values of precipitation.
The MSE is 1215.74 and the RSME 34.87.

## Summary
We fitted a LASSO model for predicting the mean precipitation
in the Central Amazon Basin and used a 5-fold blocked Cross Validation
approach to find the optimal level of regularization.
After training the model we evaluated its performance on a separate validation set that was not used in the training process.
The model shows predicting capabilities but misses out on higher values of the precipitation target.
Also although predicting seems to bring useful results, the choice of the LASSO coefficients seems somewhat arbitrary or at least not
interpret able in a straightforward way.

