# Lasso on cluster 3

```{r, echo = FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE, echo = FALSE)
```


```{r}
library(patchwork)
library(ggpubr)
library(raster)
library(glmnet)
library(Hmisc)
source("../code/R/helper-functions.R")
```

```{r}
path_to_model_folder <- "../results/CV-lasso/cluster-cv-lasso-og2/cluster-3/"
```


## LASSO model

We fit a LASSO model on the precipitation and SST data.
The precipitation target is the monthly mean precipitation
in the Central Amazon Basin, the SST data are monthly temperatures
over the globe.
We use a 5-fold CV approach to find an optimal lambda.
Each fold consists of 5 consecutive years of training data
followed by 2 years of test data.
In each fold we fit a LASSO model on a set of predetermined lambda 
values and choose the lambda that minimizes the MSE on the test set 
in that fold.
After determining the best lambda in each fold we choose the
lambda that minimizes the MSE over all folds and refit the model
to the complete training data. Afterwards we evaluate the fitted model on a separate validation set with 5 years length which was not included
in the training phase.

## Error plots


```{r err-bar-plot-lasso-og-cl3, fig.cap="Mean squared error of the 5-fold blocked cross validation for cluster 3. On the x-axis the lambda values on the log scale. The points in the middle represent the average MSE for the respective lambda, the errorbars give the MSE +/- one standard deviation. The dotted line shows the lambda for which minimum MSE was obtained."}
err_bars_plot <- readRDS(paste0(path_to_model_folder, "/err-mat-plots/err-bars-plot.rds"))
err_bars_plot
```


```{r err-fold-lasso-og-cl3, fig.cap="MSE of the CV for the different lambda values on the a log scale. The red dotted line shows the lambda for which minimum MSE was obtained."}
p1 <- readRDS(paste0(path_to_model_folder, "err-mat-plots/err-plot-fold-1.rds"))
p2 <- readRDS(paste0(path_to_model_folder, "err-mat-plots/err-plot-fold-2.rds"))
p3 <- readRDS(paste0(path_to_model_folder, "err-mat-plots/err-plot-fold-3.rds"))
p4 <- readRDS(paste0(path_to_model_folder, "err-mat-plots/err-plot-fold-4.rds"))
p5 <- readRDS(paste0(path_to_model_folder, "err-mat-plots/err-plot-fold-5.rds"))

p1 + p2 + p3 + p4 + p5
```

The trajectories in \@ref(fig:err-fold-lasso-og-cl3) are quite similar in fold 1
4 and 5, but quite different from fold 2 and 3.
The chosen values of regularisation differ between the folds.

```{r}
err_mat_list <- list(p1,p2,p3,p4,p5)
err_mat_resc_list <- replot_err_mat(err_mat_list)
```

```{r err-fold-lasso-og-replot-cl3, fig.cap="MSE of the CV for the different lambda values on the a log scale computed for cluster 3. The red dotted line shows the lambda for which minimum MSE was obtained. See @\ref(fig:err-folg-lasso-og), but this time the y-axis has the same range for all plots."}
err_mat_resc_list[[1]] + err_mat_resc_list[[2]] + err_mat_resc_list[[3]] + err_mat_resc_list[[4]] + err_mat_resc_list[[5]]
```

Below we can see the minimum MSE for each fold.
```{r}
#apply(err_mat, 2, min)
```
And which lambda in the lambda vector resulted in the lowest prediction error on the folds´ test set.

```{r}
#apply(err_mat, 2, function(x) which.min(x))
```



## Coefficient plots

```{r coef-plot-lasso-og-cl3, fig.cap="Coefficient map plot for the different folds. Longitude and Latitude on the x and y-axis respectively. Positive values are coloured in blue, negative values in red."}
pc1 <- readRDS(paste0(path_to_model_folder, "coef-plots/coef-plot-fold-1.rds"))
pc2 <- readRDS(paste0(path_to_model_folder, "coef-plots/coef-plot-fold-2.rds"))
pc3 <- readRDS(paste0(path_to_model_folder, "coef-plots/coef-plot-fold-3.rds"))
pc4 <- readRDS(paste0(path_to_model_folder, "coef-plots/coef-plot-fold-4.rds"))
pc5 <- readRDS(paste0(path_to_model_folder, "coef-plots/coef-plot-fold-5.rds"))



l <- list(pc1,pc2,pc3,pc4,pc5)
ggarrange(plotlist = l, ncol = 3, nrow = 2,
         common.legend = TRUE)
```


## Inspect predictions from each fold

Following we inspect the folds precipitation time series and
the predictions made by the model.


```{r}
pred_plot_1 <- readRDS(paste0(path_to_model_folder, "/pred-plots/pred-plot-fold-1.rds"))
pred_plot_1
```
```{r}
pred_plot_2 <- readRDS(paste0(path_to_model_folder, "/pred-plots/pred-plot-fold-2.rds"))
pred_plot_2
```
```{r}
pred_plot_3 <- readRDS(paste0(path_to_model_folder, "/pred-plots/pred-plot-fold-3.rds"))
pred_plot_3
```
```{r}
pred_plot_4 <- readRDS(paste0(path_to_model_folder, "/pred-plots/pred-plot-fold-4.rds"))
pred_plot_4
```
```{r}
pred_plot_5 <- readRDS(paste0(path_to_model_folder, "/pred-plots/pred-plot-fold-5.rds"))
pred_plot_5
```

## Inspect predictions from best CV-lambda

We choose now the best lambda with the 1SE rule and fit 
the model anew to predict the time series we held out
as validation data




```{r}
full_preds <- readRDS(paste0(path_to_model_folder, "/pred-plots/pred-plot-full.rds"))
full_preds
mse_full <- get_mse_from_pred_plot(full_preds)
mse_full
```


## Summary
