### deseas lasso
```{r}
library(patchwork)
library(ggpubr)
library(raster)
library(glmnet)
library(Hmisc)
source("../code/R/helper-functions.R")
```

```{r}
path_to_model_folder <- "../results/CV-lasso/test-deseas-lasso/"
```

```{r}
err_mat <- readRDS(paste0(path_to_model_folder, "/err-mat.rds"))
lambdas <- readRDS(paste0(path_to_model_folder, "/lambda-vec.rds"))
wm <- which.min(apply(err_mat, 1, mean))
full_model <- readRDS(paste0(path_to_model_folder, "full-model.rds"))
intercept <- round(full_model$a0[wm],2)
lambda <- round(lambdas[wm],2)
rm(full_model)
```

```{r err-bar-plot-lasso-deseas, fig.cap="Mean squared error of the 5-fold blocked cross validation for a range of lambda values on the log scale. The points in the middle represent the average MSE for the respective lambda, the errorbars give the MSE +/- one standard deviation. The dotted line shows the lambda for which minimum MSE was obtained."}
err_bars_plot <- readRDS(paste0(path_to_model_folder, "/err-mat-plots/err-bars-plot.rds"))
err_bars_plot
# a <- ggplot_build(err_bars_plot)
```

Initially with decreasing regularization the MSE decreases as well
before it reaches its minimum and stays fairly constant afterwards (\@ref:(err-bar-plot-lasso-deseas)).
In the decreasing area although the size of the error bar increases
which we will inspect further in the error plots for each fold.

```{r err-fold-lasso-deseas, fig.cap="MSE of the CV for the different lambda values on the a log scale. The red dotted line shows the lambda for which minimum MSE was obtained."}
p1 <- readRDS(paste0(path_to_model_folder, "/err-mat-plots/err-plot-fold-1.rds"))
p2 <- readRDS(paste0(path_to_model_folder, "/err-mat-plots/err-plot-fold-2.rds"))
p3 <- readRDS(paste0(path_to_model_folder, "/err-mat-plots/err-plot-fold-3.rds"))
p4 <- readRDS(paste0(path_to_model_folder, "/err-mat-plots/err-plot-fold-4.rds"))
p5 <- readRDS(paste0(path_to_model_folder, "/err-mat-plots/err-plot-fold-5.rds"))

p1 + p2 + p3 + p4 + p5
```

Fold 4 causes the errorbars to increase in range, since it has a local maximum
around l$log(\lambda)$) 1.25. Fold 5 shows a similar behaviour but the
size of the local maxima is smaller relatively to the other MSE values for this fold. Apart from that folds 1,2,3,5 show similar trajectories although the range
of their MSE values differ (\@ref(fig:err-fold-lasso-deseas)).

```{r}
pred_plot_1 <- readRDS(paste0(path_to_model_folder, "/pred-plots/pred-plot-fold-1.rds"))
pred_plot_1 <- pred_plot_1 + ylab("Precipitation fold 1")
mse_1 <- get_mse_from_pred_plot(pred_plot_1)

pred_plot_2 <- readRDS(paste0(path_to_model_folder, "/pred-plots/pred-plot-fold-2.rds"))
pred_plot_2 <- pred_plot_2 + ylab("Precipitation fold 2")
mse_2 <- get_mse_from_pred_plot(pred_plot_2)

pred_plot_3 <- readRDS(paste0(path_to_model_folder, "/pred-plots/pred-plot-fold-3.rds"))
pred_plot_3 <- pred_plot_3 + ylab("Precipitation fold 3")
mse_3 <- get_mse_from_pred_plot(pred_plot_3)

pred_plot_4 <- readRDS(paste0(path_to_model_folder, "/pred-plots/pred-plot-fold-4.rds"))
pred_plot_4 <- pred_plot_4 + ylab("Precipitation fold 4")
mse_4 <- get_mse_from_pred_plot(pred_plot_4)

pred_plot_5 <- readRDS(paste0(path_to_model_folder, "/pred-plots/pred-plot-fold-5.rds"))
pred_plot_5 <- pred_plot_5 + ylab("Precipitation fold 5")
mse_5 <- get_mse_from_pred_plot(pred_plot_5)

pred_plot_list <- list(pred_plot_1,pred_plot_2,pred_plot_3,pred_plot_4,pred_plot_5)
```

```{r pred-plot-fold-lasso-deseas, fig.cap="Precipitation prediction and target values in the test set in each fold. Predictions in red and target values in black."}
pred_plot_1 + pred_plot_2 + pred_plot_3 + pred_plot_4 +
  pred_plot_5
```

In \@ref(fig:pred-plot-fold-lasso-deseas) we observer that the predictions
are not too different from the other models so far, with the exception of
fold 3 and 4, where the fold models underestimate mor strongly the precipitation test values.


```{r pred-plot-full-lasso-deseas, fi.cap="Precipitation prediction and target values in the validation set. Predictions in red and target values in black. The model was fitted on the full CV data with the lambda value that minimised the average MSE"}
full_preds <- readRDS(paste0(path_to_model_folder, "/pred-plots/pred-plot-full.rds"))
full_preds
mse_full <- get_mse_from_pred_plot(full_preds)
mse_full
```

The final predictions appear to be less smooth than in the other models so far
(lasso with and without standardization), but generally as well
underestimate the 



```{r coef-plot-full-lasso-deseas, fig.cap=paste("Coefficient plot of the full lasso model with fitted intercept of", intercept)}
coef_full <- readRDS(paste0(path_to_model_folder,
                     "coef-plots/coef-plot-full.rds"))
coef_full + theme(legend.position = "bottom")
```