# Fused Lasso Regression

## The fused lasso

```{r}
# TODO add statistical learning with sparsity to references.
```

As expected and seen in the results, the LASSO choose single
SST regions as predictors as opposed to whole regions.

```{r}
# TODO make reference to achieved results here
```

```{r}
# TODO cite general LASSO properties here
# how it acts in situations where coefficients are grouped
# or highly correlated.
```

```{r}
# TODO add plot of the graph given by our helper functions.
```

```{r}
# TODO add graphic of fig.2 fused lasso
```

The fused LASSO solves the problem:

$$\min_{\beta} 1/2 \sum_{i=1}^n(y_i - x_i^T\beta_i)^2 + \lambda \sum_{i,j \in E} |\beta_i - \beta_j| + \gamma \cdot \lambda \sum_{i=1}^p|\beta_i|  ,$$
with $x_i$ being the ith row of the predictor matrix and 
E is the edge set of an underlying graph.
$\hat{\beta}$ is computed as a function of $\lambda$, with fixed $\gamma$. $\gamma$ > 0 introduces sparsity to the fused solution, with higher values placing more priority on the sparsity of the solution. 


